A formal approach in train regulation

### 2025-06-05 paper/main.tex
images

This contribution finds its origin in a cyber-warefare we are organising for the IT students at the University of Lorraine, see~\cite{CHE}. In the game-play, we have both IT infrastructures and physical systems, with connections between them. On the physical side, we built a small train network that tries to be a good model of real systems.  Our system manages customer informations, train regulation and finally rail infrastructure. Attackers can hit any of the three parts: disfiguring the site, changing a dock or blocking a light. Defenders will protect their system.

@@

From an IT point of view, our system is shaped as shown on the right. We find two databases, one for the customers, one for the infrastructure, switches are controlled by a raspberry (TURNOUT) running an OpenPLC program, signals are controlled by an other raspberry and both communicate via modbus. The SIGNAL automaton read mouvement sensors (not all links shown in the figure). Next to that, we have a third machine, the regulator: it is in charge of updating train routes (as required by the customer IT),  of the regulation of turnouts and finally it may lock some signal (turn it to "red"). All the network connections are done via a single switch, connections from REGULATOR to TURNOUT and SIGNAL are done via modbus while the customer IT and the regulator talks  are based on TCP/IP.  To activate train mouvements, we use the DCC protocol via an Arduino, but this part is strictly related to the fact we have a model train.

@@

 Routes are not computed in advance. The schedule is renewed every "day" and we introduce some randomness in the process: some trains are not the usual ones, some trains departure/destination may change, and so on. In our model, we distinguish freight and  passengers routes giving priority to the latter.

>>>>

This contribution finds its origin in a cyber-warefare we are organising for the IT students at the University of Lorraine, see~\cite{CHE}. In the game-play, we have both IT infrastructures and physical systems, with connections between them. On the physical side, we built a small train network that tries to be a good model of real systems.  Our system manages customer informations, train regulation and finally rail infrastructure. Attackers can hit any of the three parts: disfiguring the site, changing a platform or blocking a signal. Defenders will protect their system, typically via a strict network policy.

@@

From an IT point of view, our system is shaped as shown on the right. Switches are controlled by a raspberry (TURNOUT) running an OpenPLC program, signals are controlled by an other raspberry and both communicate via modbus. The SIGNAL automaton reads mouvement sensors outputs (not all links shown in the figure). Next to that, we have a third machine, the regulator: it is in charge of updating train routes (as required by the customer IT),  of the regulation of turnouts and finally it may lock some signal (turn it to "red"). All the network connections are done via a single switch, connections from REGULATOR to TURNOUT and SIGNAL are done via modbus while the customer IT and the regulator talks  are based on TCP/IP.  To activate train mouvements, we use the DCC protocol via an Arduino, but this part is strictly related to the fact we have a model train.

@@

\begin{wrapfigure}[8]{r}{0.28\textwidth}
 \begin{minipage}{0.28\textwidth}
        \centering
        \vspace{-1mm}
\includegraphics[height=50mm]{maquette.png}
    \end{minipage}
\end{wrapfigure}

Final words on the N-scale model. Our circuit has 8 blocks, 5 switches. Each block is bidirectional with a signal at each end.  We operate  three trains in parallel. Due to the size of the model, we reduced our signals  to "red" and "green" status, the train being too close one from the other. Furthermore, in case of a red signal, we make the hypothesis that the train will be able to stop before it reaches the end of the block. In practice, we stop the trains in the middle.

Routes are not computed in advance. The schedule is renewed every "day"\footnote{We use a time scale factor: one of our "day" corresponds to 20 minutes in real time.}  and we introduce some randomness in the process: some trains are inserted, some trains departure/destination may change, and so on. Finally, in our model, we distinguish freight and  passengers routes giving priority to the latter. Routes are computed each time a train as finished its duty.

 Now, we can enter in more details about how the train regulation works. First, for trains, we have a simple policy.  The regulator will ask the train to start in some direction (Forward/Backward). Then, the train will continue until a signal stops it and then restart when the signal turns green. It will do that up to its arrival point. On the other side, the regulator will, depending on the position of the trains, change switches position, lock/unlock some signal, ask for a train departure. All these operations are done by what we call an order. Trains Orders have the shape "Start direction Until position". Regulator orders are: XXX

  We do not make hypotheses on the durations of travels, but once a train is started and not stopped at some point, it will reach in finite time the next block.
